---
title: "Boxplots"
output: html_notebook
---


```{r}
library(ggplot2)
# Data Wrangling ----------------------------------------------------------
setwd("/Users/siddia/Box/Plots")
#write the file for direct 
df.QN.long<-read.delim("//Users/siddia/Box/Plots/Duplicated.tsv")
#write.table(df.QN.long,file="/Users/siddia/Box/Plots/Duplicated.tsv",sep="\t",row.names=FALSE)

#verify
length(unique(df.QN.long$GenesSet))
#15 Clusters 

df.QN.long$GenesSet<- factor(df.QN.long$GenesSet, levels=c(   "D", "E", "F", "G", "H", "I", "J", "K", "L", "M", "N", "O", "A", "B", "C"))




# Plot Visualization ------------------------------------------------------


#pdf(file="/Users/siddia/Google Drive/1-FluVaccinesStudyManuscript/Figure2/Figure2b/Output/Figure2b.pdf",
#    width=4.6,
#    height=8.2)


# Box plot
ggplot(df.QN.long, aes(x=Rsq, y=GenesSet))+
  geom_boxplot(fill="grey",  alpha=0.5)+                                        # boxplot fill color
  theme_set(theme_bw(base_size  = 6, base_family = "Times"))+                             # base font family and size
  theme(text=element_text(color = "black"),                                               # only axis titles font color
        axis.text = element_text(color = "black"))+                                       # axis  font color
  ylab("Genes Sets")+                                                                     # y-axis labels 
  xlab (expression(paste("R"^"2")))+                                                      # x-axis label
  theme(axis.title = element_text(size = 8))                                              # Axis titles font size


#dev.off()


#saving a plot Command 

```

#TLDR: 
prefer to use personalized themes and just replace their elements like below

```{r}
#reference: https://rpubs.com/mclaire19/ggplot2-custom-themes
#https://ggplot2.tidyverse.org/reference/theme_get.html

#when we use theme () meaning pass anything in that we overrise current thme 
# we can use it with other themes 

personal_theme <- function() {
    font<-"Arial"
    theme_bw()%+replace%
  
  
  theme(

   #text elements
      plot.title = element_text(             #title
                   family = font,            #set font family
                   size = 20,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = 0,                #left align
                   vjust = 2),               #raise slightly
      
      plot.subtitle = element_text(          #subtitle
                   family = font,            #font family
                   size = 14),               #font size
      
      plot.caption = element_text(           #caption
                   family = font,            #font family
                   size = 9,                 #font size
                   hjust = 1),               #right align
      
      axis.title = element_text(             #axis titles
                   family = font,            #font family
                   size = 10),               #font size
      
      axis.text = element_text(              #axis text
                   family = font,            #axis famuly
                   size = 9),                #font size
      
      axis.text.x = element_text(            #margin for axis text
                    margin=margin(5, b = 10))
      
      #since the legend often requires manual tweaking 
      #based on plot content, don't define it here
    )
  
}







# Box plot
ggplot(df.QN.long, aes(x=Rsq, y=GenesSet))+
  geom_boxplot(fill="grey",  alpha=0.5)+                                        # boxplot fill color
  ylab("Genes Sets")+                                                                     # y-axis labels 
  xlab (expression(paste("R"^"2")))+   
  personal_theme()
```







### Titles Labels fonts and sizes   

```{r}

# Box plot
ggplot(df.QN.long, aes(x=Rsq, y=GenesSet))+
  geom_boxplot(fill="grey",  alpha=0.5, #coef=1e30,
               
               outlier.colour = "white",
                 outlier.shape = 24,
                 outlier.fill = "white",
                 outlier.size = 3 # not actually needed
)+                                        # boxplot fill color
  
  theme_set(theme_bw(base_size  = 6, base_family = "Times"))+                             # base font family and size
  theme(text=element_text(color = "black", size=18))  +                                        # both x and y axes  axis both together 
        # axis.title.y =element_text(color = "red", size=18),                               # for only y or x axis title 
        # axis.title.x =element_text(color = "Green", size=15),                             # for only y or x axis title 
        # axis.text = element_text(color = "red", size=10))+                                       # axis  font color
  ylab("Genes Sets")+                                                                     # y-axis labels 
  xlab (expression(paste("R"^"2")))+                                                      # x-axis label
  theme(axis.title = element_text(size = 8))                                              # Axis titles font size


```


### Fonts families 
We use theme() to tweak the display of an existing theme.

```{r}

theme_bw(base_size=12, base_family = "")

```



### Font realated things 

TLDR:We need for mac to get the font we need to use in ggplot2

library(extrafont)
font_import()
fonts()

```{r}
#https://github.com/wch/extrafont
#install.packages('extrafont')
library(extrafont)
#After intalling the package extraFont  first time 
# have to run font_import like this (it took like 5 minutes)
#do not skip this part
font_import()# import all fonts on my system
#loadfonts(device = "win")#for windows
#loadfonts()#for mac

#see hat fonts you have available with the command windowsFonts()
#windowsFonts()
fonts()#display present fonts 



```

You can find the name of a font you need for the family parameter of element_text with the following code snippet:
```{r}
names(wf[wf=="Arial"])
```



```{r}
# Box plot
ggplot(df.QN.long, aes(x=Rsq, y=GenesSet))+
  geom_boxplot(fill="grey",  alpha=0.5, coef=1e30,
  
  theme_set(theme_bw(base_size  = 6, base_family = "Arial"))+                             # base font family and size
  theme(text=element_text(color = "black", size=18))  +                                        # both x and y axes  axis both together 
        # axis.title.y =element_text(color = "red", size=18),                               # for only y or x axis title 
        # axis.title.x =element_text(color = "Green", size=15),                             # for only y or x axis title 
        # axis.text = element_text(color = "red", size=10))+                                       # axis  font color
  ylab("Genes Sets")+                                                                     # y-axis labels 
  xlab (expression(paste("R"^"2")))                                                      # x-axis label
  #theme(axis.title = element_text(size = 8))                                              # Axis titles font size

```



### problems related to Outlier 

How to deal eith their visualisations 
couple of options discussed here 

Option 1: 
use the color option to matcvh the outliers with background
```{r}
ggplot(df.QN.long, aes(x=Rsq, y=GenesSet))+
geom_boxplot(fill="grey",  alpha=0.5, #coef=1e30,
               outlier.colour = "white",
                 #outlier.shape = 24,
                 outlier.fill = "white",
                 #outlier.size = 3 # not actually needed
)


```


second use the high limit then default 1.5 in coef option to do it.

Caveat is that it extends the whiskers and not rescale your plot.

It force the whiskers to extend to the max and min values, you can tweak the coef argument. Default value for coef is 1.5 (i.e. default length of the whiskers is 1.5 times the IQR).
```{r}
ggplot(df.QN.long, aes(x=Rsq, y=GenesSet))+
geom_boxplot(fill="grey",  alpha=0.5, coef=1e30)

ggplot(df.QN.long, aes(x=Rsq, y=GenesSet))+
geom_boxplot(fill="grey",  alpha=0.5, coef=1)
```




#References
For outlers related :https://stackoverflow.com/questions/5677885/ignore-outliers-in-ggplot2-boxplot
Font of axis titles etc: https://statisticsglobe.com/increase-font-size-in-plot-in-r
